---
title: "Red Wine Exploratory Data Analysis by R"
author: "*Anouar Zbaida*"
date: "*nov 28, 2017*"
output: "html_document"
---

# Introduction

In this project we will explores the univariate, bivariate, & multivariate
relationships between variables using exploratory data analysis (EDA)
techniques in R. To do so we are going to use a 
[tidy data](http://vita.had.co.nz/papers/tidy-data.pdf) 
that is created - using red wine samples - on 2009 by P.cortez and al, 
the dataset is related to variants of the Portuguese 
["Vinho Verde"](https://en.wikipedia.org/wiki/Vinho_Verde) wine.

### About the dataset
This dataset is public available for research. The details are described in
[Cortez et al., 2009].

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
Modeling wine preferences by data mining from physicochemical properties.
In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at:

- [Elsevier](http://dx.doi.org/10.1016/j.dss.2009.05.016)
- [Pre-press (pdf)](http://www3.dsi.uminho.pt/pcortez/winequality09.pdf)
- [bib](http://www3.dsi.uminho.pt/pcortez/dss09.bib)


### Variable description

- **Fixed acidity**: most acids involved with wine or fixed or nonvolatile 
(do not evaporate readily)
- **Volatile acidity**: the amount of acetic acid in wine, which at too high
of levels can lead to an unpleasant, vinegar taste
- **Citric acid**: found in small quantities, citric acid can add 'freshness'
and flavor to wines
- **Residual sugar**: the amount of sugar remaining after fermentation stops,
it's rare to find wines with less than 1 gram/liter and wines with greater
than 45 grams/liter are considered sweet
- **Chlorides**: the amount of salt in the wine
- **Free sulfur dioxide**: the free form of SO2 exists in equilibrium between
molecular SO2 (as a dissolved gas) and bisulfite ion; it prevents microbial
growth and the oxidation of wine
- **Total sulfur dioxide**: amount of free and bound forms of S02; in low
concentrations, SO2 is mostly undetectable in wine, but at free SO2
concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine
- **Density**: the density of water is close to that of water depending on the
percent alcohol and sugar content
- **pH**: describes how acidic or basic a wine is on a scale from 0 (very
acidic) to 14 (very basic); most wines are between 3-4 on the pH scale
- **Sulphates**: a wine additive which can contribute to sulfur dioxide gas
(S02) levels, wich acts as an antimicrobial and antioxidant
- **Alcohol**: the percent alcohol content of the wine
- **Quality**: output variable (based on sensory data, score between 0 and 10)

### Variable information

Input variables (based on physicochemical tests):

- fixed acidity (tartaric acid - g / dm^3)
- volatile acidity (acetic acid - g / dm^3)
- citric acid (g / dm^3)
- residual sugar (g / dm^3)
- chlorides (sodium chloride - g / dm^3
- free sulfur dioxide (mg / dm^3)
- total sulfur dioxide (mg / dm^3)
- density (g / cm^3)
- pH
- sulphates (potassium sulphate - g / dm3)
- alcohol (% by volume)

Output variable (based on sensory data): 

- quality (score between 0 and 10)

# Exploring Data

### Loading up libraries for project

- ggplot2 : Needed for graphs
- dplyr: Needed for data manipulation functions
- gridExtra: Allows multiple graphs in a single grid
- reshape: used for melt and cast functions (data aggregation)
- RColorBrewer: Creates nice looking color palettes especially for thematic
maps
- lattice: an implementation of Trellis graphics for R
- scales: Generic plot scaling methods
- memisc: This package collects an assortment of tools that are intended to
make work with R easier
- reshape: melt and cast functions (data aggregation)
- sandwich: Constructing sandwich covariance matrix estimators by multiplying
bread and meat matrices
- graphics:  functions for 'base' graphics

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(RColorBrewer)
library(lattice)
library(scales)
library(memisc)
library(reshape)
library(sandwich)
library(GGally)
library(graphics)
```

### Summary Statistics

```{r echo=FALSE, Load_the_Data}
# Load the Data
redwine <- read.csv('wineQualityReds.csv')
```

Let's run some basic functions to examine the structure and schema of the data
set.

```{r echo=FALSE, observations_variables}
str(redwine)
```
There are 1599 observations and 13 variables in this data set.

```{r echo=FALSE, summary}
summary(redwine)
```

- No observations have a quality greater than 8
- Quality is an ordinal categorical variable. Since we're primarily interested
in quality a new variable will be created from quality for subsquent analysis
- All other variables seem to be continuous quantities (except the
sulfur.dioxide suffixes)
- From the variable descriptions, it appears that fixed.acidity ~
volatile.acidity and free.sulfur.dioxide ~ total.sulfur.dioxide may possible
by dependent, subsets of each other

```{r echo=FALSE, quality_variable}
#Transforming Quality from an Integer to a Factor
redwine$quality <- factor(redwine$quality, ordered = T)

#Creating a new Factored Variable called 'Rating'
redwine$rating <- ifelse(redwine$quality < 5, 'bad', ifelse(
  redwine$quality < 7, 'average', 'good'))

redwine$rating <- ordered(redwine$rating,
                       levels = c('bad', 'average', 'good'))
```

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots}

# Create some functions to not write too much code #

# Find the outlier line, which is 3.5 standard deviations 
# more than the median of the variable.
positive_outlier_line <- function(variable) {
  return (geom_hline(yintercept = (median(variable) + sd(variable)*3.5),
                     alpha = 1/3, linetype = 2))
}

# Find the median to project in the histogram
median_histogram <- function(variable) {
  return(geom_vline(xintercept = median(variable), 
                    color = "red", alpha = 1/3))
}

# Create a boxplot and two histograms (normal, log10)
# The 'variable_name' is used to label the axis.

boxplot_histogram <- function(variable, variable_name) {
  return (grid.arrange(
    ggplot(aes(x = 1, y = variable), data = redwine) + 
      geom_jitter(alpha = 0.1) + 
      geom_boxplot(alpha = 0.2, color = 'blue') + 
      stat_summary(fun.y=mean, shape=1, col = 'red', geom = 'point') + 
      positive_outlier_line(variable) +
      ylab(variable_name), 
    ggplot(aes(variable), data = redwine) + 
      geom_histogram(bins=30, color = 'white') + 
      median_histogram(variable) +
      labs(x = variable_name), 
    ggplot(aes(log10(variable)), data = redwine) + 
      geom_histogram(bins=30, color = 'white') +
      labs(x = paste("log10", variable_name)),
    ncol=3))
}

```
In this section, we should perform some preliminary exploration of
our dataset. Let's run some summaries of the data and create univariate 
plots to understand the structure of the individual variables 
in our dataset.

### Quality

The goal of this studies is trying to understand which variables are
responsible for the quality of the wine. So let's start with quality variable.

```{r echo=FALSE, quality}
grid.arrange(ggplot(data = redwine, aes(x = quality)) + 
               geom_bar(width = 1, color = 'white'), 
             ggplot(data = redwine, aes(x = rating)) + 
               geom_bar(width = 1, color = 'white'),
             ncol = 2)
summary(redwine$quality)
```

- A large majority of the wines examined received ratings of 5 or 6, and very
few received 3, 4, or 8
- More than 50 % of wines rated average

### Fixed Acidity

```{r echo=FALSE, fixed.acidity}
boxplot_histogram(redwine$fixed.acidity, "Fixed Acidity")
summary(redwine$fixed.acidity)
```

- Fixed acidity has a long-tail distribution. 
- The log10 plot normalize the distribution.
- A peak around 7, a median of 7.90 and vary around 4 and 15.90

### Volatile Acidity

```{r echo=FALSE, volatile.acidity}
boxplot_histogram(redwine$volatile.acidity, "Volatile Acidity")
summary(redwine$volatile.acidity)
```

- Similar to fixed acidity, volatile acidity also has a long tail
distribution.
- Normally-distributed with peaks around 0.4 and 0.7, and a median of 0.52

### citric Acid

```{r echo=FALSE, citric.acid}
boxplot_histogram(redwine$citric.acid, "Citric Acid")
summary(redwine$citric.acid)

# length(subset(redwine, citric.acid == 0)$citric.acid)
```

- 132 wines have citric acid 0.0
- Citric.acid, did not appear to be normally-distributed with long-tail
distribution
- For the log10. Looking at this new distribution it is interesting that the
distribution has switched from a long-right-tail to a long-left-tail 
distribution
- There are alot of 0 which means a lot of wine don't has citric acid. It make
sense because citric acid is added as a freshner or flavor to wines not as a 
key ingredients in cooking wine. ([and also it is prohibited in the
EU](https://en.wikipedia.org/wiki/Acids_in_wine#Citric_acid))

### Residual Sugar

```{r echo=FALSE, residual.sugar}
boxplot_histogram(redwine$residual.sugar, "Residual Sugar")
summary(redwine$residual.sugar)
```

- Residual sugar has a very long-tail distribution with many outliers. Some of
these outliers are more than 9 standard deviations away from the median! It
will be interesting to see how these outliers affect the quality of wine. 
- In the log10 plots, the values are still very skewed, but it looks more like
a normal distribution.
- High peaks at around 2.3 with many outliers present at the higher ranges.

### Chlorides 

```{r echo=FALSE, chlorides}
boxplot_histogram(redwine$chlorides, "Chlorides")
summary(redwine$chlorides)
```

- Very similar to residual sugar, chlorides have a strong concentration of
points around the median, as well as many outliers. 
- The more extreme outliers are as much as 11 standard deviations away from
the mean. 
- In the log10 plots, we can see that the data still has a long-tail skew, but
looks more normally distributed. 

### Free Sulphur Dioxide 

```{r echo=FALSE, free.sulfur.dioxide}
boxplot_histogram(redwine$free.sulfur.dioxide, "Free Sulphur Dioxide")
summary(redwine$free.sulfur.dioxide)
```

- For Free Sulphur Dioxide, there is a high peak at 7 but then it again
follows the same positively skewed long tailed patterns
- The median is 14 and 75% of free sulfur concentration is less than 21 but a
maximum value is of 72.
- In log10 plot, the distribution appears to be binomial and dispersed

### Total Sulfur Dioxide

```{r echo=FALSE, total.sulfur.dioxide}
boxplot_histogram(redwine$total.sulfur.dioxide, "Total Sulfur Dioxide")
summary(redwine$total.sulfur.dioxide)
```

- Total sulfur dioxide distribution not that dispersed, as the free sulfur
dioxide and its interquartile range does not look quite as large. 
- In log10 plot, looks normally distributed.
- The median is 38. 75% of the wines have a concentration below 62.

### Density

```{r echo=FALSE, density}
boxplot_histogram(redwine$density, "Density")
summary(redwine$density)
```

- Density varies from 0.99 to 1 only. Its a very little variation. Most of the
wines have around 0.997 density value.
- The distribution is normal.

### PH

```{r echo=FALSE, pH}
boxplot_histogram(redwine$pH, "PH")
summary(redwine$pH)
```

- Normal looking distribution, with most of the pH values falling between 3.1
and 3.5 and a median of 3.310

### Sulphates

```{r echo=FALSE, sulphates}
boxplot_histogram(redwine$sulphates, "sulphates")
summary(redwine$sulphates)
```

- The sulfates distribution is skewed to the right and has many big outliers
- In log10 plot sulphates is much more normally distributed
- Most of the values of sulphate are around 0.6

### Alcohol

```{r echo=FALSE, alcohol}
boxplot_histogram(redwine$alcohol, "Alcohol")
summary(redwine$alcohol)
```

- Most wines have less than 11% alcohol.
- The alcohol has distribution skewed to the right, some peaks and a median of 10.2.
- Only few values of alcohol amount lie below 9 and above 13.

# Univariate Analysis

### What is the structure of your dataset?

The Red Wine Dataset had 1599 observations with 13 variables. All the variables
are numerical, expect for quality, which is an categorical. The categorical
variable indicate the quality of the wine, at least 3 wine experts rated the
quality of each wine, and rated it between 0 (bad) and 10 (excellent).

### What is/are the main feature(s) of interest in your dataset?

By univariate analysis, we could notice :

- 82.5% of the wines are rated either a 5 or 6
- Most wines have an alcohol content of less than 11%
- 34% fixed.acidity values between [7,8] and 78% Citric.acid values below 0.5
- pH and density are the most normally distributed features 
- Residual sugar, chlorides, sulphates all have big outliers

We want to analyse the quality, so quality is the main feature of interest.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Unsure how the multiple acidity variables relate to each other and quality.
We notice that Citric.acid is the only variable with observations equal to zero.
And we think sulphates, chlorides may help to analyse the quality of wine as
well.

### Did you create any new variables from existing variables in the dataset?

Yes, we changed quality to an ordered factor and created a new variable called
"rating"" to classifying wines as bad, average or good based on quality.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

- Citric acid has a unique distribution as compared to the other numeric
variables. It almost has a rectangular shape apart from a few outliers
- More than 80% rathed between 5 and 6, so this less informations will make it
hard to go out with  a good result
- Tidy adjustments not made to data
- Change the quality type into an ordered feature

# Bivariate Plots Section

In this section we will analyse our red wine dataset with 2 variables at a
time. starting with a correlation graph may help us. 

```{r echo=FALSE, Correlation_graph}
# make the quality numeric so it doesnt' get ignored in the coorelation graph
redwine$quality<- as.numeric(redwine$quality)
ggcorr(redwine %>% 
         dplyr::select(-X) %>%
         dplyr::select(-rating), # we could easily use this redwine[, 2:13]
  hjust = 0.60,
  size = 3,
  label = TRUE,
  label_alpha = TRUE,
  geom = "circle",
  max_size = 7,
  size = 3,
  hjust = 0.75,
  angle = 0,
  palette = "RdBu")+
  ggplot2::labs(title = "Correlation graph between variables")

```

- This plot shows that quality is strongly correlated positively with alcohol 
and sulfates, and negatively with volatile.acidity
- Residual.sugar on the other hand is not correlated at all with 
quality
- Density and fixed.acidity are two features with strong positive correlation.
- There is also an expected strong negative correlation between pH and 
fixed/citric.acid, and another negative correlation between alcohol and density.
- Volatile.acidity has a positive correlation with pH! (We know that when 
decreasing pH, acidity increases

**To get a quick snapshot of how the variables affect quality we will use \
boxplots**

```{r echo=FALSE, boxplot_bivariate}
# Create a function to generate boxplots
get_bivariate_boxplot <- function(x, y, ylab) {
  return(ggplot(aes(factor(x), y), data = redwine) +
           geom_jitter( alpha = .3)  +
           geom_boxplot( alpha = .5,color = 'blue')+
           stat_summary(fun.y=mean, shape=1, col = 'red', 
                        geom = 'point', size = 1) +
           ylab(ylab))
}

# Create boxplots and arrange them
grid.arrange(get_bivariate_boxplot(redwine$rating, redwine$fixed.acidity,
                                   'fixed acidity'),
             get_bivariate_boxplot(redwine$rating, redwine$volatile.acidity,
                                   'volatile acidity'),
             get_bivariate_boxplot(redwine$rating, redwine$citric.acid,
                                   'citric acid'),
             get_bivariate_boxplot(redwine$rating, redwine$residual.sugar,
                                   'residual sugar'),
             get_bivariate_boxplot(redwine$rating, redwine$chlorides,
                                   'chlorides'),
             get_bivariate_boxplot(redwine$rating, redwine$free.sulfur.dioxide,
                                   'free sulf. dioxide'),
             get_bivariate_boxplot(redwine$rating,
                                   redwine$total.sulfur.dioxide,
                                   'total sulf. dioxide'),
             get_bivariate_boxplot(redwine$rating, redwine$density,
                                   'density'),
             get_bivariate_boxplot(redwine$rating, redwine$pH,
                                   'pH'),
             get_bivariate_boxplot(redwine$rating, redwine$sulphates,
                                   'sulphates'),
             get_bivariate_boxplot(redwine$rating, redwine$alcohol,
                                   'alcohol'),
             ncol = 3)

```

A 'good' wine generally has these trends:

- Higher fixed acidity and citric acid, lower volatile acidity
- Lower pH 
- Higher sulphates
- Higher alcohol

**Let's see now the relation between different features**

```{r echo=FALSE, acid_relation1}

ggplot(data = redwine, aes(x = fixed.acidity, y = citric.acid)) +
  geom_point() +
  geom_smooth(method = "lm")
cor.test(redwine$fixed.acidity, redwine$citric.acid)

```

- Citric acid and fixed acidity have a strong positive correlation of 0.67
- Citric acid has a weak positive correlation of 0.23 with quality while fixed
acidity has a very weak positive correlation of 0.12.

```{r echo=FALSE, acid_relation2}

ggplot(data = redwine, aes(x = volatile.acidity, y = citric.acid)) +
  geom_point() +
  geom_smooth(method = "lm")
cor.test(redwine$volatile.acidity, redwine$citric.acid)

```

- Citric acid and volatile acidity have a negative correlation of -0.55

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Correlation plot helped to understand the correlation among different features. 
Quality is strongly correlated positively with alcohol and sulfates, and
negatively with volatile acidity. Good wines have lower pH values, which also
goes with having more fixed and citric acid.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

- Citric acid and fixed acidity have a strong positive correlation of 0.7, while
citrict acid and volatile acidity have a moderate negative correlation of -0.6
- Density and fixed acidity are two features with strong positive correlation of
0.7
- Negative correlation between alcohol and density
- An expected strong negative correlation between pH and fixed and citric acid
- A surprising positive correlation between pH and volatile acidity, since a
higher pH value means less acidity, but a higher volatile acidity means more
acidity.

### What was the strongest relationship you found?

```{r echo=FALSE, relationship_between_features}
# function to return correlation
cor_test <- function(x, y) {
  return(cor(as.numeric(x), as.numeric(y)))
}
# calculate normal correlations
correlations <- c(
  cor_test(redwine$fixed.acidity, redwine$quality),
  cor_test(redwine$volatile.acidity, redwine$quality),
  cor_test(redwine$citric.acid, redwine$quality),
  cor_test(redwine$residual.sugar, redwine$quality),
  cor_test(redwine$chlorides, redwine$quality),
  cor_test(redwine$free.sulfur.dioxide, redwine$quality),
  cor_test(redwine$total.sulfur.dioxide, redwine$quality),
  cor_test(redwine$density, redwine$quality),
  cor_test(redwine$pH, redwine$quality),
  cor_test(redwine$sulphates, redwine$quality),
  cor_test(redwine$alcohol, redwine$quality))

names(correlations) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'residual.sugar', 'chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'sulphates', 'alcohol')

# calculate correlation on log10
correlations_log10 <- c(
  cor_test(log10(redwine$fixed.acidity), redwine$quality),
  cor_test(log10(redwine$volatile.acidity), redwine$quality),
  cor_test(log10(redwine$citric.acid), redwine$quality),
  cor_test(log10(redwine$residual.sugar), redwine$quality),
  cor_test(log10(redwine$chlorides), redwine$quality),
  cor_test(log10(redwine$free.sulfur.dioxide), redwine$quality),
  cor_test(log10(redwine$total.sulfur.dioxide), redwine$quality),
  cor_test(log10(redwine$density), redwine$quality),
  cor_test(log10(redwine$pH), redwine$quality),
  cor_test(log10(redwine$sulphates), redwine$quality),
  cor_test(log10(redwine$alcohol), redwine$quality))

names(correlations_log10) <- c('fixed.acidity', 'volatile.acidity', 'citric.acid',
                         'residual.sugar', 'chlordies', 'free.sulfur.dioxide',
                         'total.sulfur.dioxide', 'density', 'pH',
                         'sulphates', 'alcohol')
```

- Normal correlations between quality and other features

```{r echo=FALSE, normal_correlations}
correlations
```

- Correlations log10 between quality and other features

```{r echo=FALSE, log10_correlations}
correlations_log10
```

We can say that the following variables have relatively higher correlations to
wine quality:

- alcohol
- volatile acidity
- sulphates (log10)
- citric acid

Between other features, the strongest relationship appeared to be pH and fixed
acidity, which had a nagtive correlation of -0.683.

# Multivariate Plots Section

In this section we are going to create a few multivariate plots to investigate
more complex interactions between the variables that are most related to quality
of the wine. Then we willcreate some mathematical models.

```{r echo=FALSE, Multivariate_Plots}
redwine$quality <- factor(redwine$quality, ordered = T)

grid.arrange(
  ggplot(data = redwine, aes(x = volatile.acidity, y = alcohol)) + 
    facet_wrap(~rating) + 
    geom_point(aes(color = quality)), 
  ggplot(data = redwine, aes(x = log10(sulphates), y = alcohol)) + 
    facet_wrap(~rating) +
    geom_point(aes(color = quality)),
  ggplot(data = redwine, aes(x = citric.acid, y = alcohol)) + 
    facet_wrap(~rating) + 
    geom_point(aes(color = quality)),
  ggplot(data = redwine, aes(x = volatile.acidity, y = log10(sulphates))) +
    facet_wrap(~rating) + 
    geom_point(aes(color = quality)),
  ggplot(data = redwine, aes(x = citric.acid, y = log10(sulphates))) + 
    facet_wrap(~rating) + 
    geom_point(aes(color = quality)), 
  ggplot(data = redwine, aes(x = citric.acid, y = volatile.acidity)) +
    facet_wrap(~rating) + 
    geom_point(aes(color = quality)),
  ncol = 2)

```

These scatterplots were crowded since more than 80% have average quality, so we
faceted by rating. It is now more clear that higher quality wines tend to be
higher in alcohol, citric acid and sulphates. On the other hand, higher quality
wines tend to be lower in volatile acidity.

### Linear models

```{r echo=FALSE, Linear_models}
m1 <- lm(as.numeric(quality) ~ alcohol, data = subset(redwine, citric.acid == 0))
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + I(log10(sulphates)))
m4 <- update(m3, ~ . + citric.acid + chlorides + density + pH)

mtable(m1, m2, m3, m4)
summary(m3)
```

It is difficult to predict the quality of wine, given this dataset. We are able to
predict 36.2% of what makes a good wine good. After adding chlorides, density and
pH to our model we could only improve the prediction with 2%.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

High alcohol contribute to good quality of wine, adding sulphates or citric acid
will influence the quality of wine positively, while adding volatile acid
influences the quality of wine negatively.

### Were there any interesting or surprising interactions between features?

Correlation plot was showing that citric acid influence the quality of wine but
from the plots above, we can observe that citric acid alone doen't influence the
quality that much.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

We created a linear model starting from alcohol adding volatile.acidity,
sulphates, chlorides, density and pH, we could only predict 38.2% of the quality.
We are not entirely surprised by this limitations given that the correlation
between any variable and quality was not very high, and 80% of the dataset was in
average.

------

# Final Plots and Summary

### Plot One:  Quality of wine
```{r echo=FALSE, Plot_One}
ggplot(data = redwine, aes(x = quality)) + 
  geom_bar(width = 1, color = 'white', fill = 'blue') +
  geom_vline(xintercept = median(as.numeric(redwine$quality)), color = "red") +
  labs(x = "Quality of wine", 
       y = "Number of simples", 
       title = "Distribution of wine Quality")

summary(redwine$quality)
```

This graph explains that 82.5 % of wines in dataset are of quality 5 & 6. As we
have to find influence of other metrics on quality of wine such a limited dataset
make it difficult to understand what makes good wine.


### Plot Two: Effect of Alcohol
```{r echo=FALSE, Plot_Two}
ggplot(data = redwine, aes(x = factor(quality), y = alcohol, fill = rating)) +
  geom_jitter(alpha = .3)  +
  geom_boxplot(alpha = .5, color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               size = 1) +
  geom_hline(yintercept = 11, linetype="dashed") +
  labs(x = "Quality",
       y = "Alcohol (% by Volume)",
       title = "Effect of Alcohol on Wine Quality")

cor.test(redwine$alcohol, as.numeric(redwine$quality))
by(redwine$alcohol, redwine$rating, summary)

```

Alcohol has the strongest correlation with quality. As the alcoholic content
increases, typically the quality of wine does as well. 75% of good wine contain
above 11% of alcohol, while 75% of average and bad quality ones have a alcohol
percent bellow 11%.

### Plot Three: Alcohol and volatile acidity
```{r echo=FALSE, Plot_Three}
ggplot(data = subset(redwine, rating != 'average'),
       aes(x = volatile.acidity, y = alcohol, color = rating)) + 
  geom_jitter(size = 2, alpha = 1/2) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, size = 1) +
  labs(x = "Volatile Acidity",
       y = "Alcohol (% by Volume)",
       title = "Alcohol vs. Volatile Acidity on Wine Quality")
```

*We subsetted the data to take off the 'average' wines.* Hight volatile acidity -
with few exceptions - made wine quality down. *We notice that the lines make is
more clear to see the relationship alcohol and volatile acidity by rating.* 
**High quality wine have a combination of high alcohol content and low volatile 
acidity**

------

# Reflection

Through this exploratory data analysis of red wine dataset, We can assume how
domain knowledge is helpful during the process. We could reveals the key factor
affecting the quality of wine, mainly: alcohol, sulphates, and volatile acidity
even tough we had a limited data of 1599 observations. In that dataset 82% of the
wines are of avergae quality betwen 5 & 6. If we could have a dataset of more
observations and an uniform quality of wines. Then we will be able to perform a
better analysis.

The linear model which we have used for the prediction is not a good model as its
coefficient of determination was very low. For the further study we would try
differnt modeling technic; it might give more accuracy.

*Enjoy the code 
[here](https://github.com/anoru/Red-Wine-Exploratory-Data-Analysis-by-R)*

------

# Opening to more reseach for the reader

[Receiving Operating Characteristic, or ROC](https://en.wikipedia.org/wiki/Receiver_operating_characteristic), 
is a visual way for inspecting the performance of a binary classifier (0/1). This
is just a test, and our binary classifier for this dataset is if the quality more
than 5 (1), or less than 5 (0). We are comparing the rate at which our classifier
is making correct predictions (True Positives or TP) and the rate at which your
classifier is making false alarms (False Positives or FP). We will use a
generalized linear models (Just to change the linear model)

for more informations check this : 
[ROC Curves in Python and R](http://blog.yhat.com/posts/roc-curves.html)
```{r echo=FALSE, ROCR}
redwine$acceptable <- redwine$quality > 5
is_test <- runif(nrow(redwine)) > 0.75
training <- redwine[is_test == FALSE,]
testing <- redwine[is_test == TRUE,]

summary(fit <- glm(as.numeric(quality) ~ alcohol + volatile.acidity + 
                     I(log10(sulphates)) + citric.acid + chlorides + 
                     density + pH , data = training))                   

library(ROCR)

prob <- predict(fit, newdata = testing, type = "response")
pred <- prediction(prob, testing$acceptable)
perf <- performance(pred, measure = "tpr", x.measure = "fpr")

auc <- performance(pred, measure = "auc")
auc <- auc@y.values[[1]]

roc.data <- data.frame(fpr=unlist(perf@x.values),
                       tpr=unlist(perf@y.values),
                       model="GLM")
ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
    geom_ribbon(alpha=0.2) +
    geom_line(aes(y=tpr)) +
    ggtitle(paste0("ROC Curve w/ AUC=", auc)) +
  geom_abline(linetype='dashed')
```